---
title: "DataAnalysis"
output: html_document
authour: Kevindeep Saroya
date: 23/Jan/2020
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
library("dplyr")

# Introduction

+ Some functions not included in base rstudio will be required to perform the calculations
+ Download "dplyr" and "ggplot2" using the following command:
```{r eval=FALSE}
install.packages("dplyr")
instal.packages("ggplot2")
```
+ Load "dplyr" and "ggplot2" into the the script to get access to the functions that will allow us to manipulate the data and create plots
```{r}
library("dplyr")
library("ggplot2")
```


# Data Generation

### Loading our Data

+ Assigning our data to an object in R
```{r}
MData<-read.csv("/Users/kevinsaroys/Desktop/Biol 432/assignment1/Measurements.csv")
print(MData)
```

# Data Manipulation

### Shortening column names

+ To alter the original column names, a "dplyr" command use rename() function

```{r}
MData<-rename(MData,Spe=Species)
MData<-rename(MData,Lln=Limb.Length)
MData<-rename(MData,Lwd=Limb.Width)
MData<-rename(MData,unL=unitsL)
MData<-rename(MData,unW=unitsW)
```

+ verify the name change function worked using names() functions

```{r echo=TRUE}
names(MData)
```

+ Updated data set

```{r}
MData
```

+ Changing measurement units to match one common unit


```{r}
MData$unL<-c("cm")
MData$unW<-c("cm")
```


+ Updated MData list with units changed
```{r}
print(MData)
```

### Creating the Volume column

```{r}
MData<-mutate(MData, Volume= Lln*Lwd)
MData
```

## Plotting our generated data

+ Width and length will be plotted using the function qplot()

```{r evalu=TRUE}
Plot1<-qplot(data=MData,x=Lln,y=Lwd,ylab="Limb Length (cm)",xlab="Limb Width (cm)",size=I(4),color=I("red"))
print(Plot1)
```

+ Next we will plot Width and Length for species that have above-average volume
+ First we must calculate the average volume of the species which will be used to sort our data later
```{r evalu=TRUE}
mean(MData$Volume)

```

+ Value obtained was 94.8 (rounded)
+ Using this value as a filter, we will using the 'pipe' functions of 'dplyr' to narrow down our data set and make a plot for above-average species only 
+ MData is listed at the beginning, every subsequent line our data.frame is being sent through the functions until we get an output

```{r}
MData %>% 
  select(Spe,Lln,Lwd,Volume) %>% 
  filter(Volume>=94.8)
```

+ Plottting this data

```{r}
Plot2<-MData %>% 
  select(Spe,Lln,Lwd,Volume) %>% 
  filter(Volume>=94.8) 
qplot(data=Plot2,x=Lln,y=Lwd,xlab="Limb Length (cm)", ylab="Limb Width (cm)",color=I("red"),fill=I("blue"),size=I(4))
```

+ Creating a table for species with below-average volume only

```{r}
Plot3<-MData %>% 
  select(Spe,Volume) %>% 
  filter(Volume<=94.8) 
print(Plot3)
```

+ Now we must group the species into 2 seperate and arbitrary groups

+ First grouping will be done based on taxon, each species will be assigned a 1 or 2 to group them into seperate taxons
+ Second grouping will be done based on whether the animal can fly, will be grouped into YES or NO
+ Then we will summarize the data using the summarize() function to find mean and standard deviation across the groups

```{r}
MData<-mutate(MData, Taxon=sample(x=c("One","Two"),prob=c(0.5,0.5),size=10,replace=TRUE))
MData<-mutate(MData, Flight=sample(x=c("YES","NO"),prob=c(0.5,0.5),size=10,replace=TRUE))
print(MData)
```





## Generating Summary data for our data.frame

+ Summarize the mean and SD of Length, Width and Volume for each of the groups, and present in table form.

+ First group is Taxon, included are the summaries of mean and sd for Length, Width, and Volume
```{r}
MData %>% 
  select(Spe,Lln,Lwd,Volume,Taxon) %>% 
  filter(Taxon=="One") %>% 
  summarize(MeanLln=mean(Lln),MeanLwd=mean(Lwd),MeanVOL=mean(Volume),SDLln=sd(Lln),SDLwd=sd(Lwd),SDVOL=sd(Volume))


```
+ The other group in Taxon will be see as:
```{r}
MData %>% 
  select(Spe,Lln,Lwd,Volume,Taxon) %>% 
  filter(Taxon=="Two") %>% 
  summarize(MeanLln=mean(Lln),MeanLwd=mean(Lwd),MeanVOL=mean(Volume),SDLln=sd(Lln),SDLwd=sd(Lwd),SDVOL=sd(Volume))
```

+ Next, the summaries for the Flight group who are "YES"
```{r}
MData %>% 
  select(Spe,Lln,Lwd,Volume,Flight) %>% 
  filter(Flight=="YES") %>% 
  summarize(MeanLln=mean(Lln),MeanLwd=mean(Lwd),MeanVOL=mean(Volume),SDLln=sd(Lln),SDLwd=sd(Lwd),SDVOL=sd(Volume))
```

+ Summaries for the Flight group who are "NO"
```{r}
MData %>% 
  select(Spe,Lln,Lwd,Volume,Flight) %>% 
  filter(Flight=="NO") %>% 
  summarize(MeanLln=mean(Lln),MeanLwd=mean(Lwd),MeanVOL=mean(Volume),SDLln=sd(Lln),SDLwd=sd(Lwd),SDVOL=sd(Volume))
```











